Concurrent Programming with CXL

### 2020-05-17 book/paper.tex
Latest

\chapter{Naive Attempts at Implementing Critical Sections}

\begin{figure}
\begin{code}
\verbatiminput{UpLock.cxl}
\end{code}
\caption{Naive implementation of a lock.}
\label{fig:uplock}
\end{figure}

\begin{figure}
\begin{code}
\verbatiminput{UpFlags.cxl}
\end{code}
\caption{Naive use of flags to solve mutual exclusion.}
\label{fig:upflags}
\end{figure}

\begin{figure}
\begin{code}
\verbatiminput{UpTurn.cxl}
\end{code}
\caption{Naive use of turn variable to solve mutual exclusion.}
\label{fig:upturn}
\end{figure}

Figure~\ref{fig:uplock} presents a solution based on a \emph{lock}.
The idea is that the lock is a baton that at most one process can own at
a time.
Initially the lock is not owned, indicated by lock being \texttt{False}.
To enter the critical section, a process waits until lock is \texttt{False}
and then sets it to \texttt{True} to indicate that the lock has been taken.
The process then executes the critical section.  Finally the process
releases the lock by setting it back to \texttt{False}.

Unfortunately, if we run the program, we find that the assertion still
fails.  Diagnosing the problem, we see that the \texttt{lock} variable
suffers from the same problem as the \texttt{count} variable: operations
on it consist of several instructions are divisible.  It is thus possible
for both processes to believe the lock is available and obtain the lock
at the same time.

Figure~\ref{fig:upflags} presents a solution based on each process having
a flag indicating that it is trying to enter the critical section.
After setting the flag, the process waits until the other process
(\texttt{1 - self}) is not trying to enter the critical section.
If we run this program, the assertion does not fail.  In fact, this
solution does prevent both processes being in the critical section at
the same time.
However, it has a different problem called livelock: if both try to
enter the critical section at the same time, they may end up waiting for
one another.

>>>>

\chapter{Na\"{\i}ve Attempts at Implementing Critical Sections}

\begin{figure}
\begin{code}
\verbatiminput{UpLock.cxl}
\end{code}
\caption{Na\"{\i}ve implementation of a lock.}
\label{fig:uplock}
\end{figure}

\begin{figure}
\begin{code}
\verbatiminput{UpFlags.cxl}
\end{code}
\caption{Na\"{\i}ve use of flags to solve mutual exclusion.}
\label{fig:upflags}
\end{figure}

\begin{figure}
\begin{code}
\verbatiminput{UpTurn.cxl}
\end{code}
\caption{Na\"{\i}ve use of turn variable to solve mutual exclusion.}
\label{fig:upturn}
\end{figure}

Figure~\ref{fig:uplock} presents a solution based on a na\"{\i}ve implementation
of a \emph{lock}.
The idea is that the lock is a baton that at most one process can own at
a time.
Initially the lock is not owned, indicated by lock being \texttt{False}.
To enter the critical section, a process waits until lock is \texttt{False}
and then sets it to \texttt{True} to indicate that the lock has been taken.
The process then executes the critical section.  Finally the process
releases the lock by setting it back to \texttt{False}.

Unfortunately, if we run the program, we find that the assertion still
fails.  Diagnosing the problem, we see that the \texttt{lock} variable
suffers from the same problem as the \texttt{count} variable
in Figure~\ref{fig:inc}: operations
on it consist of several instructions are divisible.  It is thus possible
for both processes to believe the lock is available and obtain the lock
at the same time.

Figure~\ref{fig:upflags} presents a solution based on each process having
a flag indicating that it is trying to enter the critical section.
A process can write its own flag and read the flag of its peer.
After setting its flag, the process waits until the other process
(\texttt{1 - self}) is not trying to enter the critical section.
If we run this program, the assertion does not fail.  In fact, this
solution does prevent both processes being in the critical section at
the same time.
However, it has a different problem called livelock: if both try to
enter the critical section at the same time, they may end up waiting for
one another.

The final na\"{\i}ve solution is based on a variable called \texttt{turn}
that alternates between 0 and 1.  When \texttt{turn = $i$}, process~$i$ can
enter the critical section, while process $1-i$ has to wait.  When done,
process~$i$ sets \texttt{turn} to $1-i$ to give the other process an
opportunity to enter.
The solution has the nice properties that it satifies mutual exclution and,
moreover, processes 0 and 1 alternate entering the critical section.
The problem with the solution is
that it violates the \emph{progress} property:
if process~$i$ terminates instead of entering the critical section when it
is process~$i$'s turn, process $1-i$ ends up waiting indefinitely for its
turn.

\chapter{Peterson's Algorithm}

\begin{figure}
\begin{code}
\verbatiminput{../Peterson.cxl}
\end{code}
\caption{Peterson's Algorithm}
\label{fig:peterson}
\end{figure}

\chapter{Spinlock}
